---
title: "Experimenting with acsr "
output: html_notebook
---

ACSR is a wrapper for the acs package.

https://github.com/sdaza/acsr
http://sdaza.com/r/2016/07/06/acsr/ 

"The acsr package helps extracting variables and computing statistics using the America Community Survey and Decennial US Census. It was created for the Applied Population Laboratory (APL) at the University of Wisconsin-Madison."

```{r packages & apikey, message=FALSE, warning=FALSE}

devtools::install_github("sdaza/acsr")   
library(acsr)
library(tidyverse)

api.key.install(key=mykey)

```

Lets try out the package and get some small scale working prototype


```{r kicking the acsr wheels}

# working example from acsr.  Single aggregation
sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
       varname = "mynewvar",
       method = "prop",
       level = "tract",
       state = "CA",
       county = 075,
       tract = as.list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
       one.zero = TRUE,
       combine = TRUE)

# working example from acsr.  using first 2 nhood track sets.  very verbose method first
combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("tract", "tract"),  # if two aggregations need list that defines each aggregation (all 2 are at the tract level) 
    state = list("CA", "CA"), # for each aggregation have to define all three levels: state, county, and tract
    county = list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
                  list(075, 075, 075, 075, 075, 075)),
    tract = list(list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
                 list(25200, 25200, 25402, 25403, 25300, 25100)),
    one.zero = TRUE,
    combine.names = c("CA1", "CA2")) # give them names

# test with 3
level <- list("tract", "tract", "tract")
state <- list("CA", "CA", "CA")
county <- list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
              list(075, 075, 075, 075, 075, 075),
              list(075,075))
tract <-  list(
            list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
             list(25200, 25200, 25402, 25403, 25300, 25100),
             list(25402,25403))
combine.names  <-  c("CA1", "CA2", "CA3")

# working example using the 'test with 3' vars from above
combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = level,
    state = state,
    county = county,
    tract = tract,
    one.zero = TRUE,
    combine.names = combine.names)

# works. ok so I need to automate process from portal data that feeds the level, state, county, tract, and combine.name vars

```


Bring in census tract and nhood and SD crosswalk from open data portal
- goal is can just update the dataset in 2020 any things will propogate
- we wil need to create a SD/nhood crosswalk like we have for nhood (Ill bring it up with Jason)

Analysis Neighborhood 2010 Census Tracks - https://data.sfgov.org/d/bwbp-wk3r

```{r crosswalks, message=FALSE, warning=FALSE}
nhoodxwalk <- read_csv("https://data.sfgov.org/api/views/nq42-rde6/rows.csv?accessType=DOWNLOAD")

# selecting and renaming the fields I want.  Creating county column for use later
nhoodxwalk <- 
  nhoodxwalk %>% 
  mutate(TRACTCE10 = as.numeric(TRACTCE10),
         county = 075) %>% 
  select(nhood = NHOOD,
         tract = TRACTCE10,
         county)

nhoodxwalk



#convert to list.  Which is the format acsr prefers

nhoodxwalk <- 
  nhoodxwalk %>% 
  group_by(nhood) %>% 
  summarise(tract = list(tract), 
            county = list(county),
            state = "CA",
            level = "tract")


nhoodxwalk

tract <-  lapply(nhoodxwalk$tract, as.list)
county <-  lapply(nhoodxwalk$county, as.list)
state <- as.list(nhoodxwalk$state)
level <- as.list(nhoodxwalk$level)
combine.names <- nhoodxwalk$nhood

# survey year
endyear <- 2015



```

from our base variables can we get it to run.
-yes. but its hella slow. need to investigate acsdata function that predownloads the base tables.

```{r test estimates}
#counts population.  match acs. although does not calculate cv.
combine.output("b01003_001",
    varname = "var",
    method = "aggregation",
    level = level,
    state = state,
    county = county,
    tract = tract,
    one.zero = TRUE,
    endyear = endyear,
    combine.names = combine.names)

# percent
combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "pct",
    method = "prop",
    level = level,
    state = state,
    county = county,
    tract = tract,
    one.zero = TRUE,
    combine.names = combine.names)



```
Now trying using the batch method described in the email.
We would essentially create a long csv file with varname, method, and formula and have R iterate through each row.


```{r batch test}

# bring in small example
data <- read_csv("moe_FS_15.csv")

#try on first row
combine.output(formula = data$formula[1],  #insert value in first row of formula column, which is a text string
       varname = data$myfield[1],
       method = data$type[1],
       level = level,
       state = state,
       county = county,
       tract = tract,
       one.zero = TRUE,
       combine.names = combine.names)

#can I precedurally store the results in a list to access later.
results <- list()

results[[1]] <- results
combine.output(formula = data$formula[1],  #insert value in first row of formula column, which is a text string
       varname = data$myfield[1],
       method = data$type[1],
       level = level,
       state = state,
       county = county,
       tract = tract,
       one.zero = TRUE,
       combine.names = combine.names)

# now try to loop through

results <- list() #empty out list

for(i in 1:nrow(data)){  # find the number of rows in the dataset. Cycle through 1 until that number substituting the number for the letter i.
  results[[i]] <- 
    combine.output(formula = data$formula[i],
                   varname = data$myfield[i],
                   method = data$type[i],
                   level = level,
                   state = state,
                   county = county,
                   tract = tract,
                   one.zero = TRUE,
                   combine.names = combine.names) 
}

results[[1]]
results[[2]]
results[[3]]
results[[4]]
results[[5]]
results[[6]]

```


and the final version looks something like this
```{r proof_of_concept}

# bring crosswalk and desired table codes
nhoodxwalk <- read_csv("https://data.sfgov.org/api/views/nq42-rde6/rows.csv?accessType=DOWNLOAD")
data <- read_csv("moe_FS_15.csv")

# selecting and renaming the fields I want.  Creating county column for use later
nhoodxwalk <- nhoodxwalk %>% 
  mutate(TRACTCE10 = as.numeric(TRACTCE10),
         county = 075) %>% 
  select(nhood = NHOOD,
         tract = TRACTCE10,
         county)

#convert to list.  Which is the format acsr prefers
nhoodxwalk <- 
  nhoodxwalk %>% 
  group_by(nhood) %>% 
  summarise(tract = list(tract), 
            county = list(county),
            state = "CA",
            level = "tract")

tract <-  lapply(nhoodxwalk$tract, as.list)
county <-  lapply(nhoodxwalk$county, as.list)
state <- as.list(nhoodxwalk$state)
level <- as.list(nhoodxwalk$level)
combine.names <- nhoodxwalk$nhood

# survey year
endyear <- 2015

# acsr magic!
results <- list() #empty out list

for(i in 1:nrow(data)){  # find the number of rows in the dataset. Cycle through 1 until that number substituting the number for the letter i.
  results[[i]] <- 
    combine.output(formula = data$formula[i],
                   varname = data$myfield[i],
                   method = data$type[i],
                   level = level,
                   state = state,
                   county = county,
                   tract = tract,
                   one.zero = TRUE,
                   combine.names = combine.names) 
}
```



```{r garbage}





results <- list()

for(i in 1:3){  # find the number of rows in the dataset. Cycle through that number substituting the number for the letter i.
results[[i]] <- combine.output(formula = data$formula[i],
               varname = data$myfield[i],
               method = data$type[i],
               level = list("tract", "tract"),  # if two aggregations need list that defines each aggregation (all 2 are at the tract level) 
               state = list("CA", "CA"), # for each aggregation have to define all three levels: state, county, and tract
               county = list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
                             list(075, 075, 075, 075, 075, 075)),
               tract = list(list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
                            list(25200, 25200, 25402, 25403, 25300, 25100)),
               one.zero = TRUE,
               combine.names = c("CA1", "CA2")) # give them names


}
results[[1]] <- combine.output(formula = data$formula[1],
               varname = data$myfield[1],
               method = data$type[1],
               level = list("tract", "tract"),  # if two aggregations need list that defines each aggregation (all 2 are at the tract level) 
               state = list("CA", "CA"), # for each aggregation have to define all three levels: state, county, and tract
               county = list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
                             list(075, 075, 075, 075, 075, 075)),
               tract = list(list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
                            list(25200, 25200, 25402, 25403, 25300, 25100)),
               one.zero = TRUE,
               combine.names = c("CA1", "CA2")) # give them names
results[[2]] <- combine.output(formula = data$formula[2],
               varname = data$myfield[2],
               method = data$type[2],
               level = list("tract", "tract"),  # if two aggregations need list that defines each aggregation (all 2 are at the tract level) 
               state = list("CA", "CA"), # for each aggregation have to define all three levels: state, county, and tract
               county = list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
                             list(075, 075, 075, 075, 075, 075)),
               tract = list(list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
                            list(25200, 25200, 25402, 25403, 25300, 25100)),
               one.zero = TRUE,
               combine.names = c("CA1", "CA2")) # give them names
results[[3]] <- combine.output(formula = data$formula[3],
               varname = data$myfield[3],
               method = data$type[3],
               level = list("tract", "tract"),  # if two aggregations need list that defines each aggregation (all 2 are at the tract level) 
               state = list("CA", "CA"), # for each aggregation have to define all three levels: state, county, and tract
               county = list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
                             list(075, 075, 075, 075, 075, 075)),
               tract = list(list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
                            list(25200, 25200, 25402, 25403, 25300, 25100)),
               one.zero = TRUE,
               combine.names = c("CA1", "CA2")) # give them names

results <- list()
results <- 0

for(i in 1:3){  # find the number of rows in the dataset. Cycle through that number substituting the number for the letter i.
results[[i]] <- 5







a <- c(1,2,3)
a <- lapply(nhoodxwalk$tract, as.list)

b <- list(1,2,3)


as.list(c(1,2,3))


# 140 state, county, tract

combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("tract", "tract"),
    state = list("CA", "CA"),
    county = list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
                  list(075, 075, 075, 075, 075, 075)),
    tract = list(list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
                 list(25200, 25200, 25402, 25403, 25300, 25100)),
    one.zero = TRUE,
    combine.names = c("CA1", "CA2"))


sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
       varname = c("mynewvar"),
       method = c("prop"),
       level = c("tract"),
       #tract = list(c()),
       state = "CA",
       county = 075,
       tract = list(nhoodxwalk$tract[[2]]),
       one.zero = TRUE,
       combine = TRUE)


sumacs("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
 varname = "test",
 method = "prop",
 level = c("block.group"),
 state = list("WI"),
 county = list(1, 141),
 tract = list(950100, 11700),
 block.group = list(1:2, 1:2),
 combine = TRUE)

combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("county", "county"),
    state = list("WI", "WI"),
    county = list( list(1:2, 3:4), list(5:6, 7:8) ),
    combine.names = c("WI+MN", "CA+OR"))

# 060 state, county, county.subdivision 

combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("county.subdivision", "county.subdivision"),
    state = list("CA", "CA"),
    county = list(001, 001),
    county.subdivision = list(list("Fremont CCD"),
                    list("Fremont CCD")),
    combine.names = c("CA1", "CA2"))

# 140 state, county, tract

combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("tract", "tract"),
    state = list("CA", "CA"),
    county = list(list(075, 075),
                  list(075, 075)),
    tract = list(list(980600, 980900),
                 list(25200, 25200)),
    combine.names = c("CA1", "CA2"))

# 060 state, county

combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("county", "county"),
    state = list("CA", "CA"),
    county = list( 
      list(075, 111, 057, 061), 
      list(113, 115, 063, 023, 021)), # nested list
    combine.names = c("CA1", "CA2"))

# 040 state 

combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("state", "state"),
    state = list( list("CA", "CA"), list("CA", "CA")), # nested list
    combine.names = c("CA1", "CA2"))

combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = list("state", "state"),
    state = list( list("WI", "MN"), list("CA", "OR")), # nested list
    combine.names = c("WI+MN", "CA+OR"))


sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
            varname = "mynewvar",
            method = "prop",
            level = "tract",
            county = 1,
            tract = 950501,
            endyear = 2013,
            one.zero = TRUE)
```

```{r}

sumacs(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
       varname = "mynewvar",
       method = "prop",
       level = "tract",
       state = "CA",
       data = a,
       county = 075,
       tract = as.list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
       one.zero = TRUE,
       combine = TRUE)

# working example from acsr.  Single aggregation
a <- acsdata(formula = "(b16004_004 + b16004_026 + b16004_048) / b16004_001",
       level = "tract",
       state = "CA",
       county = 075,
       tract = as.list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
       combine = TRUE)

acsdata(formula, 
        level = "state", 
        dataset = "acs", 
        endyear = 2014,
        span = 5, 
        state = "WI",
        county = "*", 
        tract = "*",
        combine = FALSE, 
        combine.name = "aggregate", 
        print.levels = TRUE)



# test with 3
level <- list("tract", "tract", "tract")
state <- list("CA", "CA", "CA")
county <- list(list(075, 075, 075, 075, 075, 075, 075, 075, 075, 075, 075),
              list(075, 075, 075, 075, 075, 075),
              list(075,075))
tract <-  list(
            list(980600, 980900, 61200, 23103,61000,23003,23200,23001,23400,23300,23102),
             list(25200, 25200, 25402, 25403, 25300, 25100),
             list(25402,25403))
combine.names  <-  c("CA1", "CA2", "CA3")

# working example using the 'test with 3' vars from above
combine.output("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
    varname = "mynewvar",
    method = "prop",
    level = level,
    state = state,
    county = county,
    tract = tract,
    one.zero = TRUE,
    combine.names = combine.names)

acsdata("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
        level = level,
        state = state,
        county = county,
        tract = tract,
        combine = TRUE,
        combine.name = combine.names)

```



```{r}
acsdata("(b16004_004 + b16004_026 + b16004_048 / b16004_001)", level = "county")
acsdata(c("b16004_004", "b16004_026"), level = "county")


acsdata("(b16004_004 + b16004_026 + b16004_048 / b16004_001)",
  level = c("block.group"), 
  state = list("WI"),
  county = list(1, 141),
  tract = list(950100, 11700),
  block.group = list(1:2, 1:2), 
  combine = TRUE)

```

